---
title: "Data Prep"
author: "Jin Yao"
date: "2023-07-04"
output: html_document
---

```{r}

```

------------------------------------------------------------------------

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Data

```{r}
Data <- read.csv('Final HDB price2.csv')
# Check the dimension of Data
dim(Data)
```

## Missing Data Inspection

```{r}
# Get the total number of missing data
missing_values <- colSums(is.na(Data))
missing_percentage <- (missing_values / nrow(Data)) * 100
columns_with_missing <- names(missing_percentage[missing_percentage > 0])
columns_with_missing
```

The issue is that we do not have 2023 population data for Singapore as this is concluded at the end of the year. Since population effect on HDB price is at the interest of our study, we shall not attempt to impute but to cut off our modelling period to 2022

### Remove missing population row

```{r}
# Remove rows with missing values
Data_clean <- na.omit(Data)
# Missing value final check
missing_values_final <- colSums(is.na(Data_clean))
# Check
missing_percentage_final <- (missing_values_final / nrow(Data_clean)) * 100
columns_with_missing_final <- names(missing_percentage_final[missing_percentage_final > 0])
columns_with_missing_final
```

## Perform scaling

```{r}
library(dplyr)
# Select columns
Dataset <- data.frame(select(Data_clean,date,price,town,flat_type,floor_area_sqm,storey_range,remaining_lease,children,adult,senior_citizen,total_population))
head(Dataset)
```

### Perform dummy variable creation

```{r}
# Perform dummy variable creation for town,storey range and flat type
#install.packages('fastDummies')
library('fastDummies')
Dataset1 <- dummy_cols(Dataset,select_columns = c('town', 'flat_type','storey_range'))
```

```{r}
# Unselect the town, flat_type,storey_range
Dataset1 <- subset(Dataset1, select = -c(town, flat_type, storey_range))
# Select columns to change to numeric 
selected_columns <- c("price","floor_area_sqm","remaining_lease")
#Convert selected columns to numeric
Dataset1 <- Dataset1 %>% mutate(across(all_of(selected_columns), as.numeric))
head(Dataset1)
```

### Perform scaling

Scaling is necessary when variables in a dataset have different magnitudes because it ensures that each variable has an equal contribution to the analysis or model. By applying scaling, we adjust the variables to a common scale, eliminating the dominance of variables with larger magnitudes. This allows us to compare and interpret the variables more accurately and ensure that they have a balanced impact on the response variable or any subsequent analysis performed.

```{r}
# Set the CRAN mirror
options(repos = "https://cran.rstudio.com")

# Install the "caret" package
library(caret)
# Select columns to be scaled
scaled_columns <- c("floor_area_sqm", "remaining_lease", "children","adult","senior_citizen","total_population")

# Apply min-max scaling using preProcess()
preprocessed_data <- preProcess(Dataset1[, scaled_columns], method = "range")

# Transform the selected columns using the computed scaling parameters
scaled_data <- predict(preprocessed_data, newdata = Dataset1[, scaled_columns])
```

```{r}
# Join the scaled data back to the original Dataset1
# Remove columns from Dataset1
Dataset1 <- subset(Dataset1, select = -c(floor_area_sqm, remaining_lease, children, adult, senior_citizen,date,total_population))
Dataset1 <- cbind(Dataset1[,], scaled_data)
```

```{r}
# Create boxplots for all columns except "date" and "price_adj_sqm"
boxplot(subset(Dataset1, select = c(remaining_lease, floor_area_sqm)))
boxplot(subset(Dataset1, select = c(price)))
```

## Perform Pearson Correlation Matrix

```{r}
# Select numerical variables
Collinear_dataset <- subset(Dataset1)
# Make all as numeric
# Convert columns to numeric
Collinear_dataset$price <- as.numeric(Collinear_dataset$price)
Collinear_dataset$floor_area_sqm <- as.numeric(Collinear_dataset$floor_area_sqm)
Collinear_dataset$remaining_lease <- as.numeric(Collinear_dataset$remaining_lease)
Collinear_dataset$children <- as.numeric(Collinear_dataset$children)
Collinear_dataset$adult <- as.numeric(Collinear_dataset$adult)
Collinear_dataset$senior_citizen <- as.numeric(Collinear_dataset$senior_citizen)
library("car")
library("ggplot2")
library("reshape2")

# Calculate the correlation matrix
cor_matrix <- cor(Collinear_dataset)

# Get the lower triangle of the correlation matrix
lower_tri <- cor_matrix
lower_tri[upper.tri(cor_matrix)] <- NA

# Melt the lower triangle correlation matrix
melted_cor <- melt(lower_tri, na.rm = TRUE)

# Create a ggplot for the correlation heatmap
ggplot(melted_cor, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1, 1),
                       space = "Lab", name = "Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 8, hjust = 1)) +
  coord_fixed() +
  labs(x = NULL, y = NULL) +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 4,
                               title.position = "top", title.hjust = 0)) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 2)


```

The result of the correlation matrix can be interpreted as follow: 

-   The higher the value, the most positively correlated the two variables are.

-   The closer the value to -1, the most negatively correlated they are.

If variables with VIF higher than 5 means these variables are highly collinear. The counter proposal will be to sum up all the group and use as total population.

## Perform Testing,Training and Validation data split then perform modelling with multiple linear regression without PCR

```{r}
# Unselect children, adult,senior citizen
#Dataset1 <- subset(Dataset1,select = -c(children,adult,senior_citizen))
# Set seed for reproducibility
set.seed(123)

# Split data into training, testing, and validation sets
train_indices <- sample(1:nrow(Dataset1), 0.7 * nrow(Dataset1))  # 70% for training
test_indices <- sample(setdiff(1:nrow(Dataset1), train_indices), 0.15 * nrow(Dataset1))  # 15% for testing
valid_indices <- setdiff(1:nrow(Dataset1), c(train_indices, test_indices))  # Remaining for validation

train_data <- Dataset1[train_indices, ]
test_data <- Dataset1[test_indices, ]
valid_data <- Dataset1[valid_indices, ]

# Fit the multiple linear regression model
model <- lm(price ~ ., data = train_data)

# Print the model summary
summary(model)

# Predict on the testing data
test_predictions <- predict(model, newdata = test_data)

# Predict on the validation data
valid_predictions <- predict(model, newdata = valid_data)

# Calculate the Mean Squared Error (MSE) on the testing data
mse_test <- mean((test_predictions - test_data$price_adj_sqm)^2)
cat("MSE on testing data:", mse_test, "\n")

# Calculate the Mean Squared Error (MSE) on the validation data
mse_valid <- mean((valid_predictions - valid_data$price_adj_sqm)^2)
cat("MSE on validation data:", mse_valid, "\n")

# Calculate R-squared for the testing data
r2_test <- 1 - mean((test_data$price - test_predictions)^2) / var(test_data$price)
cat("R-squared for the testing data:", r2_test, "\n")

# Calculate R-squared for the validation data
r2_valid <- 1 - mean((valid_data$price - valid_predictions)^2) / var(valid_data$price)
cat("R-squared for the validation data:", r2_valid, "\n")
```

## Perform PCA

```{r}

# Perform PCA using the final collinearity data
PCA_Dataset <- subset(Dataset1, select= -c(total_population))
# Identify non-numeric columns
non_numeric_cols <- sapply(train_data, function(x) !is.numeric(x))

# Convert non-numeric columns to numeric format
train_data[non_numeric_cols] <- lapply(train_data[non_numeric_cols], as.numeric)

# Remove observations with missing values
train_data <- na.omit(train_data)

# PCA_data
pca_data <- subset(train_data,select=-c(price))

# Perform PCA using the final collinearity data
cor_pca <- cor(pca_data)
data.pca <- princomp(cor_pca)
summary(data.pca)
```

```{r}
library('FactoMineR')
library('factoextra')
variance_plot <- fviz_eig(data.pca, addlabels = TRUE, ylim = c(0, 19), barfill = "lightblue", ncp = 57) +
  theme(panel.background = element_rect(fill = "white"),
        axis.text = element_text(size = 4))

# Set the plot size and margins
variance_plot <- variance_plot + theme(plot.margin = unit(c(50, 50, 50, 50), "pt"))


# Save a copy of the plot
ggsave("variance_plot.png", variance_plot, width = 10, height = 10, dpi = 300)

variance_plot
```

```{r}
# Method 1: direct from prcomp output
pca <- prcomp(pca_data, scale. = TRUE)
# Method 1: direct from prcomp output
PCs <- pca$x[,1:11]
PCA_DATA_TRAIN <- cbind(train_data$price,PCs)
as.data.frame(PCA_DATA_TRAIN)
model_PCR <- lm(V1~., data = as.data.frame(PCA_DATA_TRAIN))
beta0 <- model_PCR$coefficients[1]
betas <- model_PCR$coefficients[2:12]
# Predict on the testing data
non_numeric_cols <- sapply(test_data, function(x) !is.numeric(x))

```

betas

```{r}
summary(model_PCR)
```

```{r}

# Select the desired number of components
num_components <- 38  # Choose the desired number of components
selected_components <- pca$x[, 1:num_components]
# Get the loadings
loadings <- pca$rotation

# Extract loadings for selected components
selected_loadings <- loadings[, 1:num_components]

# Get the absolute values of the loadings
abs_loadings <- abs(selected_loadings)

# Find the most important features based on the maximum absolute loading for each component
important_features <- apply(abs_loadings, 2, which.max)

# Get the column names of the original dataset
original_feature_names <- colnames(Dataset1)

# Map the important feature indices to their corresponding feature names
important_feature_names <- original_feature_names[important_features]

# Print the important feature names
print(important_feature_names)


```

## PCR

I am using Principle component regression to compare to normal regression

```{r}
#install.packages('pls')
library(pls)
# Perform PCR
pcr_model <- pcr(price ~ ., data = train_data, scale = TRUE)
```

```{r}
# Calculate R-squared for the PCR model on the testing data
r_squared_test <- R2(pcr_model, data = test_data, format = "traditional")
r_squared_test

# Calculate R-squared for the PCR model on the validation data
r_squared_valid <- R2(pcr_model, data = valid_data, format = "traditional")
r_squared_valid
```

```{r}
summary(pcr_model)
```

## Choosing the number of appropriate PC.

```{r}
# RMSEP selection 
validationplot(pcr_model,val.type="RMSEP", cex.axis=0.7)
axis(side = 1, at = c(40), cex.axis=0.7)
abline(v = 25, col = "blue", lty = 3)
```

```{r}
validationplot(pcr_model, val.type="MSEP", cex.axis=0.7)
axis(side = 1, at = c(8), cex.axis=0.7)
abline(v = 25, col = "blue", lty = 3)
```

```{r}
validationplot(pcr_model, val.type="R2", cex.axis=0.7)
axis(side = 1, at = c(38), cex.axis=0.7)
abline(v = 25, col = "blue", lty = 3)
```

## Extract top 25 feature

```{r}
pcr_model_25 <- pcr(price ~ ., data = test_data, scale = TRUE, ncomp = 25)

# Get the R-squared value
r_squared <- R2(pcr_model_25)
r_squared
```

## Export Dataset1

```{r}

# Export test_data
write.csv(test_data, file = "test_data.csv", row.names = FALSE)

# Export train_data
write.csv(train_data, file = "train_data.csv", row.names = FALSE)

# Export valid_data
write.csv(valid_data, file = "valid_data.csv", row.names = FALSE)


```
